The State of Urbit:\\Eight Years After the Whitepaper

### 2024-04-27 mss.tex
Fix typo

A major part of making the namespace more usablae was to expose namespace queries over the network.  Designed in 2021 and deployed in 2023, the ``Fine''\footnote{Pronounced \emph{fee-NAY}, after French cartographer and mathematician Oronce Fine.} remote scry network protocol was added to Urbit.  The first use cases were Clay, which uses it for downloading desks, and the Gall \lstinline[style=inlinecode]{%grow} namespace, described in more detail in a later section.

@@

``Shrubbery'' refers to an architectural concept in which all user\-space data is addressible in the scry namespace.  This would include source code and application state in addition to inert file-like data.

@@

\sloppy
The Urbit galaxies have always been considered to form a senate, acting as the Shelling point for network governance.  Before going on chain, this was merely an idea – now it is a formalized reality.  Galaxies can vote on-chain, using a smart contract that has control over the other smart contracts.

@@

\begin{enumerate}
  \item  \lstinline[style=inlinecode]{|$} barbuc (produce mold builder wet gate)
  \item  \lstinline[style=inlinecode]{|@} barpat (produce wet core)
  \item  \lstinline[style=inlinecode]{$|} bucbar (produce structure to satisfy validator)
  \item  \lstinline[style=inlinecode]{$<} bucgal (filter a mold, exclude)
  \item  \lstinline[style=inlinecode]{$>} bucgar (filter a mold, match)
  \item  \lstinline[style=inlinecode]{$:} bucmic (enter manual value mode)
  \item  \lstinline[style=inlinecode]{$~} bucsig (define default custom type value)
  \item  \lstinline[style=inlinecode]{#/} haxfas (produce typed path)
  \item  \lstinline[style=inlinecode]{+$} lusbuc (type constructor arm)
  \item  \lstinline[style=inlinecode]{+*} lustar (door alias)
  \item  \lstinline[style=inlinecode]{;<} micgal (monadic do notation)
  \item  \lstinline[style=inlinecode]{^~} tissig (compose many expressions)
  \item  \lstinline[style=inlinecode]{?#} wuthax (match type, experimental)
  \item  \lstinline[style=inlinecode]{!<} zapgal (extract vase to mold)
\end{enumerate}

>>>>

A major part of making the namespace more usable was to expose namespace queries over the network.  Designed in 2021 and deployed in 2023, the ``Fine''\footnote{Pronounced \emph{fee-NAY}, after French cartographer and mathematician Oronce Fine.} remote scry network protocol was added to Urbit.  The first use cases were Clay, which uses it for downloading desks, and the Gall \lstinline[style=inlinecode]{%grow} namespace, described in more detail in a later section.

@@

``Shrubbery'' refers to an architectural concept in which all user\-space data is addressable in the scry namespace.  This would include source code and application state in addition to inert file-like data.

@@

\sloppy
The Urbit galaxies have always been considered to form a senate, acting as the Schelling point for network governance.  Before going on chain, this was merely an idea – now it is a formalized reality.  Galaxies can vote on-chain, using a smart contract that has control over the other smart contracts.

@@

\begin{enumerate}
  \item  \lstinline[style=inlinecode]{|$} barbuc (produce mold builder wet gate)
  \item  \lstinline[style=inlinecode]{|@} barpat (produce wet core)
  \item  \lstinline[style=inlinecode]{$|} bucbar (produce structure to satisfy validator)
  \item  \lstinline[style=inlinecode]{$<} bucgal (filter a mold, exclude)
  \item  \lstinline[style=inlinecode]{$>} bucgar (filter a mold, match)
  \item  \lstinline[style=inlinecode]{$:} buccol (produce tuple mold)
  \item  \lstinline[style=inlinecode]{$~} bucsig (define default custom type value)
  \item  \lstinline[style=inlinecode]{#/} haxfas (produce typed path)
  \item  \lstinline[style=inlinecode]{+$} lusbuc (type constructor arm)
  \item  \lstinline[style=inlinecode]{+*} lustar (door alias)
  \item  \lstinline[style=inlinecode]{;<} micgal (monadic do notation)
  \item  \lstinline[style=inlinecode]{^~} ketsig (store constant at compile time)
  \item  \lstinline[style=inlinecode]{?#} wuthax (match type, experimental)
  \item  \lstinline[style=inlinecode]{!<} zapgal (extract vase to mold)
\end{enumerate}
